The GitHub Open Source Development Process \author{Kevin Peterson\\ Department of Information Technology\\ Clinical Informatics Support Systems\\ Mayo Clinic\\ 200 1st Street\\ Rochester, MN 55905\\ \small \texttt{peterson.kevin@mayo.edu} }

### 2013-04-11 paper/main.tex
Updated paper.

A case study of the Apache Server project\cite{mockus2000case} shows that a dedicated community can produce software that rivals or exceeds commercial offerings. This study showed that even in the unstructured context of OSS, certain structures, hierarchies, and codes of conduct emerge. Contrast the Apache Server project to a successful project on GitHub. Although the spirit and intent may be the same, the tool set is drastically different. The intent of this work is to explore how GitHub facilitates this process, and also how it may cause it to evolve.

GitHub\footnote{https://github.com/} is a popular code hosting website that uses Git Source Code Management (SCM)\footnote{http://git-scm.com/}. GitHub... Git allows for development to take place in a more distributed manner than previously available in other SCM systems\cite{spinellis2012git}. Git also allows for a variety of work flows\cite{chacon2009pro}. These workflows may be tailored to the individual needs of a project. The Linux Kernel, for example, uses a Dictator and
Lieutenants Workflow\cite{platschekfloss}, which is hierarchical in nature. Other work flows may be more distributed, or resemble more traditional centralized SCM systems. Although work flow techniques aren't explicitly addressed in this work, it is important to note that work flow change may contribute to \textit{process change}.

@@

\subsection{Hypotheses}
The following hypotheses are presented in an effort to better understand the GitHub development process by analyzing data from actual GitHub projects. Each hypothesis will be backed by previous research in an attempt to show concrete ways in which the GitHub process may be changing the OSS development process. Where possible, results from previous research on the Apache Server[mockus2000case] is compared, as this work closely aligns with Hypothesis 2 and 3.

\noindent \textbf{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
In Open Source Software, a feeling of community belonging can be intrinsically motivating to developers\cite{lakhani2003hackers}. This feeling of belonging can be expressed passively, as a \emph{watch}, or actively, as a \emph{fork} of a repository. As \emph{watches} don't necessarily signal intent to contribute, a positive direction for an OSS project is to not only increase the number of \emph{watchers}, but also the number of \emph{forks}. As both signify a general general interest in the project, it is expected that there be some correlation between the two.

@@

This is, ultimately, a proof of ``Linus Law"\cite{raymond1999cathedral} in the context of GitHub. In other words, the more exposure the code gets (in terms of forks), the easier bugs will be to find and fix.

@@

\subsection{Research Questions}
To expand the analysis scope slightly, three qualitative research questions are posed. These, although backed by analyzed data where possible, are intended to show broader patterns, ideas, and motivations around the use of GitHub. Where possible, interviews\cite{begel2013social} where conducted to explore high level concepts and motivations.

\noindent \textbf{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
A smal core of developers, however, seems to be a historical fact of OSS development\cite{mockus2000case,mockus2002two,krishnamurthy2002cave}. At the extreme of this, a study of projects on the software hosting site Sourceforge\footnote{http://sourceforge.net} found that a large percentage of OSS development is done by lone developers. Is a small core of developers intrinsic to OSS development, or have the social aspects of GitHub and the distributed nature of Git changes this?

\noindent \textbf{Research Question 2: How is GitHub changing the OSS process?}\\
The social aspects of GitHub are an important part of the experience\cite{dabbish2012social}. User interactions and social pressures can drive OSS development in interesting ways. Social aspects of OSS development have existed before GitHub, with mailing lists\cite{mockus2000case}, gift giving mechanisms\cite{bergquist2008power}, and other code hosting sites. What is GitHub doing to facilitate change on OSS development, and how does it see OSS development changing moving forward?

@@

Matplotlib\cite{Hunter2007}, a 2D graphics package for Python, was used to render graphs and charts. NumPY and SciPY\cite{scipy} were used for data analysis calculations, such as correlation coefficients.

@@

\subsection{Hypotheses}
\textbf{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
A correlation coefficient of \input{watcher_forks_scatterplot-correlation} was observed between the number of Watchers and Forks of a repository. This suggests that a correlation between Watchers and Forks exists.

@@

\textbf{Hypothesis 2: As the number Repository Forks increase, the Issue Resolution Time will decrease.}\\
No significant correlation was found between the number of repository forks and issue resolution time. It is important to note, however, that issue resolution time is not necessarily related to software product quality in the context of GitHub. We can assert that for serveral reasons. \textit{First}, GitHub issues are not necessarily product defects. Issues may be feature requests. TODOs, or general support questions. \textit{second}, GitHub projects may have primary issue tracking services elsewhere, while using the GitHub issues as secondary services or not at all. \textit{Third}, Pham \textit{et al.} explore an emerging ``testing culture'' in GitHub -- something which could have sizable impacts on software quality but wouldn't necessarily be reflected in this metric \cite{phamcreating}.

@@

It is possible that other factors, such as issue prioritization, may have a greater impact on issue resolution time. Research does support a relationship\cite{mockus2002two} between the stated priority and fix interval. As the GitHub issue tracker can only support prioritization by free-text \textit{tags} placed on issues, it is not feasable to programmattically gather issue priority metrics.\\

\textbf{Hypothesis 3: There will be more Issue Reporters than Committers by an order of magnitude.}\\
Data does not support an exponentially increasing number of distinct Issue Reporters as compared to distinct Committers. 
As shown in Figure\ref{fig:issue_reporters_committers_scatterplot}, the slope of the relationship between the two variables is less than one. This does not seem to completely align with findings in the Apache Server project\cite{mockus2000case}.

\begin{figure}
\includegraphics[height=3in,width=3in]{images/issue_reporters_committers_scatterplot.png}
\caption{Repostitory issue reporters and committers}
\label{fig:issue_reporters_committers_scatterplot}
\end{figure}

Figure\ref{fig:issue_reporters_committers_scatterplot}, does however suggest, however, that that the two variables may be related. The hypothesis implies that there will be a correlation between the two variables -- which is demonstrated by a reasonably strong correlation cofficient of \input{issue_reporters_committers_scatterplot-correlation}. More investigation would be necessary to further examine this finding.\\

\subsection{Research Questions}
\textbf{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
The data in figure \ref{fig:committers_percentage_pie_chart} represents commit percentage breakdown. The inner percentages represent the percentage of committers in a project that that have committed the outer percentage amount of commits to that project, or:

@@

To further analyze this research question, and to explore further the GitHub ``society'' and its impact on OSS development, an interview with GitHub staff was conducted.\\\\
\textit{Interview Q1: How is GitHub changing the way OSS is being developed?}\\

\textit{Interview Q2: Many new Software Developers are will be starting their education/training already having experience with GitHub. How will Software Engineering as a discipline (and how it is taught) change to accommodate the ideas popularized by GitHub?}\\

\textit{Interview Q3: There are many non-code artifacts being hosted on GitHub, for example, Object Management Group Specifications, Laws and Statutes, books... etc. Do you see GitHub participation expanding in the non-code areas?}\\

\textbf{Research Question 3: Can GitHub be used for more that code artifacts?}\\
The OMG\textregistered standardization process is a process by which products strive to produce an interoperability statndard by way of an industry standards consortium.\cite{omgTechProcess}
The standardization process itself aims to, at its conclusion, produce a set of model artificats containing sufficient detail such that when interpreted and implemented, produce interoperable software.
At a conceptual level, this can be thought of as a software requirements engineering process as much as a standardization process. If so,  of the five main tasks of requirements engineering, Elicitation, Analysis and Negotiation, Documentation, Validation, and Management,\cite{sommerville1998requirements} we can show that GitHub is an appropriate forum for many of them.

@@

As mentioned, the CTS2 specification made heavy use of the GitHub issue tracker, as well as another OMG specification, ServD\footnote{https://github.com/servd/servd-specification}. For open source specifications, or requirements engineering activities, GitHub has shown to be a viable platform.

@@

\section{Acknowledgments}

@@

>>>>

A case study of the Apache Server project\cite{mockus2000case} shows that a dedicated community can produce software that rivals or exceeds commercial offerings. This study showed that even in the unstructured context of OSS, certain structures, hierarchies, and codes of conduct emerge. Contrast the Apache Server project to a successful project on GitHub. Although the spirit and intent may be the same, the tool set is drastically different. The intent of this work is to explore how GitHub facilitates this process, and also how it may be causing it to evolve.

GitHub\footnote{https://github.com/} is a popular code hosting website that uses Git Source Code Management (SCM)\footnote{http://git-scm.com/}. Along with hosting software repositories, GitHub incorporates social aspects of development. GitHub users have community-visible profiles, and user actions can be tracked and followed by other community actions. This type of social integration of a user's identity and actions is unique to GitHub\cite{dabbish2012social}.

Git itself, the SCM of GitHub, lends itself to the collaborative nature of GitHub by allowing for development to take place in a more distributed manner than previously available in other SCM systems\cite{spinellis2012git}.
Git also allows for a variety of work flows\cite{chacon2009pro}. These workflows may be tailored to the individual needs of a project. The Linux Kernel, for example, uses a Dictator and Lieutenants Workflow\cite{platschekfloss}, which is hierarchical in nature. Other work flows may be more distributed, or resemble more traditional centralized SCM systems. Although work flow techniques aren't explicitly addressed in this work, it is important to note that work flow change may contribute to \textit{process change}.

@@

\subsection{Hypotheses}
The following hypotheses are presented in an effort to better understand the GitHub development process by analyzing data from actual GitHub projects. Each hypothesis will be backed by previous research in an attempt to show concrete ways in which the GitHub process may be changing the OSS development process. Where possible, results from previous research on the Apache Server[mockus2000case] is compared, as this work closely aligns with Hypothesis 2 and 3.\\

\noindent \textbf{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
In Open Source Software, a feeling of community belonging can be intrinsically motivating to developers\cite{lakhani2003hackers}. This feeling of belonging can be expressed passively, as a \emph{watch}, or actively, as a \emph{fork} of a repository. As \emph{watches} don't necessarily signal intent to contribute, a positive direction for an OSS project is to not only increase the number of \emph{watchers}, but also the number of \emph{forks}. As both signify a general general interest in the project, it is expected that there be some correlation between the two.\\

@@

This is, ultimately, a proof of ``Linus Law"\cite{raymond1999cathedral} in the context of GitHub. In other words, the more exposure the code gets (in terms of forks), the easier bugs will be to find and fix.\\

@@

\subsection{Research Questions}
To expand the analysis scope slightly, three qualitative research questions are posed. These, although backed by analyzed data where possible, are intended to show broader patterns, ideas, and motivations around the use of GitHub. Where possible, interviews\cite{begel2013social} where conducted to explore high level concepts and motivations.\\

\noindent \textbf{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
A smal core of developers, however, seems to be a historical fact of OSS development\cite{mockus2000case,mockus2002two,krishnamurthy2002cave}. At the extreme of this, a study of projects on the software hosting site Sourceforge\footnote{http://sourceforge.net} found that a large percentage of OSS development is done by lone developers. Is a small core of developers intrinsic to OSS development, or have the social aspects of GitHub and the distributed nature of Git changes this?\\

\noindent \textbf{Research Question 2: How is GitHub changing the OSS process?}\\
The social aspects of GitHub are an important part of the experience\cite{dabbish2012social}. User interactions and social pressures can drive OSS development in interesting ways. Social aspects of OSS development have existed before GitHub, with mailing lists\cite{mockus2000case}, gift giving mechanisms\cite{bergquist2008power}, and other code hosting sites. What is GitHub doing to facilitate change on OSS development, and how does it see OSS development changing moving forward?\\

@@

Matplotlib\cite{Hunter2007}, a 2D graphics package for Python, was used to render graphs and charts. NumPy and SciPy\cite{scipy} were used for data analysis calculations, such as correlation coefficients.

@@

\subsection{Hypotheses}
\noindent \textbf{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
A correlation coefficient of \input{watcher_forks_scatterplot-correlation}}\unskip was observed between the number of Watchers and Forks of a repository. This suggests that a correlation between Watchers and Forks exists.

@@

\noindent \textbf{Hypothesis 2: As the number Repository Forks increase, the Issue Resolution Time will decrease.}\\
No significant correlation was found between the number of repository forks and issue resolution time. It is important to note, however, that issue resolution time is not necessarily related to software product quality in the context of GitHub. We can assert that for serveral reasons. \textit{First}, GitHub issues are not necessarily product defects. Issues may be feature requests. TODOs, or general support questions. \textit{second}, GitHub projects may have primary issue tracking services elsewhere, while using the GitHub issues as secondary services or not at all. \textit{Third}, Pham \textit{et al.} explore an emerging ``testing culture'' in GitHub -- something which could have sizable impacts on software quality but wouldn't necessarily be reflected in this metric \cite{phamcreating}.

@@

It is possible that other factors, such as issue prioritization, may have a greater impact on issue resolution time. Research does support a relationship\cite{mockus2002two} between the stated priority and fix interval. As the GitHub issue tracker can only support prioritization by free-text \textit{tags} placed on issues, it is not feasible to programmatically gather issue priority metrics.\\

\noindent \textbf{Hypothesis 3: There will be more Issue Reporters than Committers by an order of magnitude.}\\
Data does not support an exponentially increasing number of distinct Issue Reporters as compared to distinct Committers. 
As shown in Figure\ref{fig:issue_reporters_committers_scatterplot}, the slope of the relationship between the two variables is less than one. This does not seem to completely align with findings in the Apache Server project\cite{mockus2000case}.

\begin{figure}
\includegraphics[height=3in,width=3in]{images/issue_reporters_committers_scatterplot.png}
\caption{Repository issue reporters and committers}
\label{fig:issue_reporters_committers_scatterplot}
\end{figure}

Figure\ref{fig:issue_reporters_committers_scatterplot}, does however suggest, however, that that the two variables may be related. The hypothesis implies that there will be a correlation between the two variables -- which is demonstrated by a reasonably strong correlation cofficient of \input{issue_reporters_committers_scatterplot-correlation}}\unskip. More investigation would be necessary to further examine this finding.\\

\subsection{Research Questions}
\noindent \textbf{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
The data in figure \ref{fig:committers_percentage_pie_chart} represents commit percentage breakdown. The inner percentages represent the percentage of committers in a project that that have committed the outer percentage amount of commits to that project, or:

@@

\noindent \textbf{Research Question 3: Can GitHub be used for more that code artifacts?}\\
The OMG\textregistered standardization process is a process by which products strive to produce an interoperability statndard by way of an industry standards consortium.\cite{omgTechProcess}
The standardization process itself aims to, at its conclusion, produce a set of model artifacts containing sufficient detail such that when interpreted and implemented, produce interoperable software.
At a conceptual level, this can be thought of as a software requirements engineering process as much as a standardization process. If so,  of the five main tasks of requirements engineering, Elicitation, Analysis and Negotiation, Documentation, Validation, and Management,\cite{sommerville1998requirements} we can show that GitHub is an appropriate forum for many of them.

@@

As mentioned, the CTS2 specification made heavy use of the GitHub issue tracker, as well as another OMG specification, ServD.\footnote{https://github.com/servd/servd-specification} For open source specifications, or requirements engineering activities, GitHub has shown to be a viable platform.

@@

\section{Acknowledgments}
We thank Gilberto Fragoso, Ph.D. and Paul Alexander for insight and ideas, Harold Solbrig, Craig Stancl, and Dick Hedger for guidance, and Rick Kiefer for his review.

@@

\input{interview}
